---
title: "Exercise 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Graphical overview of the data and summaries of the variables

```{r echo=FALSE}
library(GGally)
library(ggplot2)
library(corrplot)
library(dplyr)
library(tidyr)
```

```{r echo=TRUE}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep  =",", header = T)
str(human)
summary(human)
```

```{r echo=TRUE}
ggpairs(human)
cor(human)%>% corrplot
```

*There are 155 observations and 8 variables. You can find more information about the variables here: https://raw.githubusercontent.com/TuomoNieminen/Helsinki-Open-Data-Science/master/datasets/human_meta.txt*
*Many of the variables seem to be normally distributed while for example Mat.Mor and GNI are not. There is a strong negative correlation between the variables 'Life.Exp' and 'Mat.Mor' and strong positive correlation between the variables 'Life.Exp' and 'Edu.Exp'*

## Perform principal component analysis (PCA)
```{r echo=TRUE}
pcahuman <- prcomp(human)
summarypcahuman <- summary(pcahuman)
summarypcahuman
```
*PC1 seems to explain nearly all of the variance.*
```{r echo=TRUE}
biplot(pcahuman, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

## Standardize the variables and repeat the above analysis
```{r echo=TRUE}
humanstd <- scale(human)
pcahumanstd <- prcomp(humanstd)
s_humanstd <- summary(pcahumanstd)
s_humanstd
pca_pr <- round(100*s_humanstd$importance[2, ], digits = 1)
pca_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")
biplot(pcahumanstd, choices = 1:2, cex = c(0.6, 1), col = c("grey40", "deeppink2"), xlab = pca_lab[1], ylab = pca_lab[2])
```

*Unfortunately, I ran out of time this week and couldn't finish the exercise.*